Ring of Fire Camera Detections Report

### 2023-02-23 analysis/paper/paper.qmd
updates

---
title: "Ring of Fire Camera Detections Report"
author:
  - Valentin Lucet:
      correspondence: "yes"
      email: valentin.lucet@gmail.com
      orcid: "0000-0003-0268-818X"
      institute: wlu
  - Frances Stewart:
      institute: wlu
      orcid: "0000-0001-9344-8346"
institute:
  - wlu:
      name: Wilfrid Laurier University
      # address: 23 Science Street, Eureka, Mississippi, USA
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/landscape-ecology.csl" # Insert path for the bib-style
abstract: |
  Abstract TBW
# keywords: |
#   keyword 1; keyword 2; keyword 3
# highlights: |
#   These are the highlights. 
---

@@

if (!file.exists("analysis/data/derived_data/annotations.rds")){
  source("scripts/1_preprocess_annotations.R")
} else {
  anns <- readRDS("analysis/data/derived_data/annotations.rds")
}

@@

dets_tb_fmt <- prettyNum(dets_tb, big.mark = ",")
```

@@

Using a cutt-off value for confidence threshold of 0.1, this left us with a total of __`r dets_tb_cat_sum_fmt`__ detections to sort through (see @fig-detections-histogram for an histogram of the detection scores, grouped by detection types). The totals for each category are: __`r dets_tb_cat_fmt[1]`__ animals (or __`r dets_tb_cat_percent[1]` %__); __`r dets_tb_cat_fmt[2]`__ persons (or __`r dets_tb_cat_percent[2]` %__); and __`r dets_tb_cat_fmt[3]`__ vehicles (or __`r dets_tb_cat_percent[3]` %__).

@@

- Deployable online. This was to allow for for multiple taggers
- Flexible tagging interface. Few platforms allow for a full customization of the tagging interface, down to the smallest details.
- Allowing to tag multiple bounding boxes per image. Most platforms only allow for tagging the entire image, and not for portions of it (e.g. bounding boxes of detections produced by MegaDetector). This is necessary for when we will crop our images to train our neural network model. 
- Optionally, open source, to allow for maximum reproducibility of the process.

@@

Out of the 189,982 animal detections, 65,338 have been processed on the platform (34.39 %) for a total of 54,677 images (1.96 % of the total amount of images, or 4.24 % of the set of images with at least one detection, or 19.44 % of the total amount of images uploaded to the platform). This does not include the empty images also viewed as part of a set, bringing that amount to 92,826 “detections” and 82,165 images, (2.94 % of the total amount of images, or 29.21 % of the total amount of images uploaded to the platform).

Out of the 65,338 detections, 30,621 were false positives (46,87 %), and 33,237 were true positives (50.86 %), to which 1,480 false negatives were added (manually added bounding boxes, 2.26 %).

- Percentage of predictions changed (no human)

@@

@@

@@

>>>>

---
title: "Ring of Fire Camera Detections Report"
author:
  - Valentin Lucet:
      # correspondence: "yes"
      email: valentin.lucet@gmail.com
      orcid: "0000-0003-0268-818X"
      institute: wlu
  - Frances Stewart:
      institute: wlu
      orcid: "0000-0001-9344-8346"
institute:
  - wlu:
      name: Wilfrid Laurier University
      # address: 23 Science Street, Eureka, Mississippi, USA
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template_edited.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/landscape-ecology.csl" # Insert path for the bib-style
abstract: |
  Abstract TBW
# keywords: |
#   keyword 1; keyword 2; keyword 3
# highlights: |
#   These are the highlights. 
---

@@

if (!file.exists("analysis/data/derived_data/annotations.rds") |
    !file.exists("analysis/data/derived_data/annotations_wide.rds")){
  source("scripts/1_preprocess_annotations.R")
} else {
  anns <- readRDS("analysis/data/derived_data/annotations.rds")
  anns_wide <- readRDS("analysis/data/derived_data/annotations_wide.rds")
}

@@

dets_tb_fmt <- prettyNum(dets_tb, big.mark = ",")

@@

Using a cut-off value for confidence threshold of 0.1, this left us with a total of __`r dets_tb_cat_sum_fmt`__ detections to sort through (see @fig-detections-histogram for an histogram of the detection scores, grouped by detection types). The totals for each category are: __`r dets_tb_cat_fmt[1]`__ animals (or __`r dets_tb_cat_percent[1]` %__); __`r dets_tb_cat_fmt[2]`__ persons (or __`r dets_tb_cat_percent[2]` %__); and __`r dets_tb_cat_fmt[3]`__ vehicles (or __`r dets_tb_cat_percent[3]` %__).

@@

- *Deployable online*. This was to allow for for multiple taggers
- *Flexible tagging interface*. Few platforms allow for a full customization of the tagging interface, down to the smallest details.
- *Allowing to tag multiple bounding boxes per image*. Most platforms only allow for tagging the entire image, and not for portions of it (e.g. bounding boxes of detections produced by MegaDetector). This is necessary for when we will crop our images to train our neural network model. 
- *Optionally, open source,* to allow for maximum reproducibility of the process.

@@

# Number of images viewed
viewed <- length(unique(anns_wide$image_id))
viewed_fmt <- prettyNum(viewed, big.mark = ",")

# Percentage of image viewed (in the total)
viewed_pct <- viewed/length(unique(dets$image_id))*100

# Percentage of image viewed from what was uploaded to the platform
viewed_pct_up <- length(unique(anns_wide$image_id))/uploaded*100

# Filter for non empty images
anns_wide_noempty <- anns_wide |>
  dplyr::filter(!is.na(to_name))

# Number of processed annotations
processed <- nrow(anns_wide_noempty)
processed_fmt <- prettyNum(processed, big.mark = ",")

# Percentage of animal annotations processed
processed_pct <- processed/dets_tb_cat[1]*100

# Number of nonempty images viewed
processed_img <- length(unique(anns_wide_noempty$image_id))
processed_img_fmt <- prettyNum(processed_img, big.mark = ",")

# Percentage of nonempty images viewed (in the total)
processed_img_pct <- processed_img/length(unique(dets$image_id))*100

# Percentage of nonempty images viewed (in the detected images > 0.1)
processed_img_up_pct <- processed_img/uploaded*100
```

Out of the __`r dets_tb_cat_fmt[1]`__ animal detections, __`r processed_fmt`__ have been processed on the platform (or __`r processed_pct` %__) for a total of __`r processed_img_fmt`__ images (i.e. __`r processed_img_pct` %__ of the total amount of images, or __`r processed_img_up_pct` %__ of the total amount of images uploaded to the platform). This does not include the empty images also viewed as part of a set, bringing that amount to 92,826 “detections” and __`r viewed_fmt`__, (i.e. __`r viewed_pct` %__ of the total amount of images, or __`r viewed_pct_up` %__ of the total amount of images uploaded to the platform).

```{r}
#| label: mans
#| echo: false

# Deleted predictions
dels <- anns_wide_noempty |> dplyr::filter(is.na(species))
dels_nb <- nrow(dels)
dels_nb_fmt <- prettyNum(dels_nb, big.mark = ",")
dels_nb_pct <- dels_nb/nrow(anns_wide_noempty)*100

# Should be no manuals here
stopifnot((anns_wide_noempty |> dplyr::filter(is.na(species)) 
           |> dplyr::filter(manual)) == 0)

# Actual annotations
true_anns <- anns_wide_noempty |> dplyr::filter(!is.na(species))
true_anns_nb <- nrow(true_anns)
# true_anns_nb_fmt <- prettyNum(true_anns_nb, big.mark = ",")
# true_anns_nb_pct <- true_anns_nb/processed*100

# Non manuals 
non_mans <- anns_wide_noempty |> dplyr::filter(!is.na(species)) |> 
  dplyr::filter(!manual)
non_mans_nb <- nrow(non_mans)
non_mans_nb_fmt <- prettyNum(non_mans_nb, big.mark = ",")
non_mans_nb_pct <- non_mans_nb/nrow(anns_wide_noempty)*100

# Mans
mans <- anns_wide_noempty |> dplyr::filter(!is.na(species)) |> 
  dplyr::filter(manual)
mans_nb <- nrow(mans)
mans_nb_fmt <- prettyNum(mans_nb, big.mark = ",")
mans_nb_pct <- mans_nb/nrow(anns_wide_noempty)*100

# adjusted
orig_tb <- anns_wide_noempty$origin |> table()
orig_tb_pct <- orig_tb/sum(orig_tb)*100
orig_tb_fmt <- prettyNum(orig_tb, big.mark = ",")
```

Out of the __`r processed_fmt`__ detections, __`r dels_nb_fmt`__ were false positives (or __`r dels_nb_pct` %__), and __`r non_mans_nb_fmt`__ were true positives (or __`r non_mans_nb_pct` %__), to which __`r mans_nb_fmt`__ false negatives were added (manually added bounding boxes, __`r mans_nb_pct` %__). Out of the __`r non_mans_nb_fmt`__ true positives, __`r orig_tb_fmt[3]`__ had to have their bounding box adjusted,

@@

```{r}
#| label: quality-control
#| eval: false
#| echo: false

# Only annotations for animals
anns_wide_animals <- anns_wide_noempty %>% 
  filter(species != 'Human') %>% 
  filter(species != 'Vehicle')

# Unknowns and bird breakdown
anns_wide_animals %>% filter(species == "Unknown") %>% pull(id) %>% 
  str_sub(1,4) %>% table() %>% sort()
anns_wide_animals %>% filter(str_detect(species,"Bird")) %>% pull(id) %>% 
  str_sub(1,4) %>% table() %>% sort()

# Join
anns_wide_animals_joined <- anns_wide_animals %>% 
  right_join(info_images, by = "image_id")
```

@@

```{r}
#| label: species
#| eval: false

# Summary by species

anns_wide_animals_table <- anns_wide_animals %>% 
  mutate(camera_id = anns_wide_animals$id %>% str_split("_") %>% 
           map(~.x[[2]]) %>% unlist()) %>% 
  dplyr::select(camera_id, species) %>% 
  filter(species != "Unknown") %>%
  table() %>% 
  `>`(0) %>% 
  `+`(0) %>% 
  as.data.frame() %>% 
  rownames_to_column("camera_id") %>% 
  arrange(camera_id)
anns_wide_animals_table_tb <- anns_wide_animals_table %>% 
  dplyr::select(-camera_id)
rownames(anns_wide_animals_table_tb) <- anns_wide_animals_table$camera_id

```

@@

```{r}
#| label: spatial
#| eval: false

RoF_CameraDeploymentLog_WLU <- 
  read_excel("data/raw/WLU_Cameras_Log_2023.xlsx") %>% 
  janitor::clean_names()

RoF_CameraDeploymentLog_WLU_old <- 
  read_excel("data/raw/RoF_CameraDeploymentLog_WLU.xlsx",
             sheet = "Sheet1") %>% 
  janitor::clean_names()

cams_sf <- RoF_CameraDeploymentLog_WLU %>% 
  filter(!is.na(latitude)) %>% 
  filter(!is.na(longitude)) %>% 
  st_as_sf(coords=c("longitude","latitude"), crs="EPSG:4326")
plot(cams_sf$geometry)

lu <- raster("data/raw/CAN_LC_2015_CAL.tif")

lu_dat <- read_csv("data/raw/lu.txt") %>% 
  mutate(cats = as.factor(code))

cams_sf_buffers <- cams_sf %>% 
  st_transform(st_crs(lu)) %>% 
  st_buffer(dist = 1000)

extr <- exact_extract(lu, cams_sf_buffers)

extr_list <- extr %>% 
  lapply(FUN = function(x){
    the_tab <- table(x$value)
    props <- the_tab/sum(the_tab)
    cats <- names(the_tab)
    df <- tibble(cats, props)
  })
names(extr_list) <- cams_sf_buffers$camera_id
extr_table <- bind_rows(extr_list, .id = "camera_id") %>% 
  mutate(props = as.double(props)) %>% 
  left_join(lu_dat, by="cats") %>% 
  dplyr::select(camera_id,category_code, props) %>% 
  # mutate(cats = paste0("cat_", cats)) %>% 
  pivot_wider(names_from = "category_code", values_from = "props") %>% 
  as.data.frame()
extr_table[is.na(extr_table)] <- 0

extr_table_sub <- extr_table %>% 
  filter(camera_id %in% anns_wide_animals_table$camera_id) %>% 
  arrange(camera_id)
extr_table_sub_tb <- extr_table_sub %>% 
  dplyr::select(-camera_id)
rownames(extr_table_sub_tb) <- extr_table_sub$camera_id

# -------------------------------------------------------------------------

# Ordination
all(rownames(anns_wide_animals_table_tb)==rownames(extr_table_sub_tb))
# plot(rda(anns_wide_animals_table_tb,extr_table_sub_tb))

test_rda <- rda(anns_wide_animals_table_tb ~ .,
                extr_table_sub_tb)

sp_scores <- scores(test_rda,
                    choices = 1:2,
                    display = "sp"
)

plot(test_rda,
     display = c("sp", "lc", "cn")
)
```
